---
title: "Peer Assignment1"
author: "Disovankiri Boung"
date: "6/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<br>
**Load Libraries**

```{r library, echo=TRUE, warning=FALSE,message=FALSE}
#load library
library(ggplot2)
library(dplyr)
library(knitr)
library(lubridate)
```

### What is mean total number of steps taken per day?

** Load the data**
```{r echo = TRUE}
#read csv file into dataframe named df
df <- read.csv("activity.csv")
head(df)
```

**1. Calculate the total number of steps taken per day**
```{r}
sum_steps_by_date <- aggregate(steps~date, df,sum,na.action = NULL )

kable(sum_steps_by_date)
```

**2.Histogram of the total number of steps taken each day**
```{r NULL,warning=FALSE,message=FALSE}
ggplot(data = sum_steps_by_date,aes(x=steps)) + geom_histogram(bins=10) +labs(x="steps",y="count",title="Histogram") + theme_bw()

```

**3. Report mean and median of number of steps taken each day**
```{r}
mean_steps_by_date <- aggregate(steps~date,df, mean,na.action = NULL)
names(mean_steps_by_date) <- c("date","mean")

median_steps_by_date <- aggregate(steps~date,df,median,na.action = NULL) 
names(median_steps_by_date) <- c("date","median")

df3 <- merge(mean_steps_by_date,median_steps_by_date)

kable(df3)
```

### What is the average daily activity pattern?

**1. Time-series plot of average steps by day**

```{r}
mean_steps_by_interval <- aggregate(steps~interval,df,mean)

ggplot(mean_steps_by_interval,aes(x=interval,y=steps)) + geom_line() + theme_bw() + labs (xlab="5 minutes interval", ylab="Average number of steps", title="Time-series of average steps by interval")
```

**2. 5 minute intervals that has maximum average steps across all day**
```{r}
max_index <- which(mean_steps_by_interval$steps==max(mean_steps_by_interval$steps))

paste("The 5-minutes interval that has the maximum average steps across days is", mean_steps_by_interval[max_index,1])
      
```

### Imputing missing values

**1. Calculate and report the total number of missing values in the datase**
```{r}
length(which(is.na(df)))
```

**2. Filling na value with mean of steps that day**

```{r}
imputed_df <- df %>%
  mutate(
    steps = case_when(
      is.na(steps) ~ mean_steps_by_interval$steps[match(df$interval, mean_steps_by_interval$interval)],      
      TRUE ~ as.numeric(steps)
    ))

any(is.na(imputed_df))
```

**3. Fill in the dataset**
```{r}
kable(head(imputed_df))
```


**4.Make a histogram of the total number of steps taken each day**
```{r}
df2_sum <- aggregate(steps~date, imputed_df,sum )
ggplot(data = df2_sum,aes(x=steps)) +
geom_histogram(bins=10) +labs(x="steps",y="count",title="Histogram") + theme_bw()
```
The total number of each day increased after we used the imputed dataset.

```{r}
df2_mean<- aggregate(steps~date,imputed_df, mean)
names(df2_mean) <- c("date","mean")

df2_median<- aggregate(steps~date,imputed_df, median)
names(df2_median) <- c("date","median")

df4 <- merge(df2_mean,df2_median)

kable(df4)
```

### Are there differences in activity patterns between weekdays and weekends?

**1. Create two factors**
```{r}

new_df <- imputed_df %>%
  mutate(
    date = ymd(date),
    day_of_week = case_when(wday(date) %in% 2:6 ~ "Weekday",
                                   wday(date) %in% c(1,7) ~ "Weekend")
  ) %>% select(-date) %>%
  group_by(interval, day_of_week) %>%
  summarise(
    steps = mean(steps)
  )

head(new_df)
```

**2. Plot**
```{r}
ggplot(new_df, aes(interval, steps)) + 
  geom_line() + 
  facet_wrap(~day_of_week, nrow = 2) +
  xlab("5-Minute intervals") + 
  ylab("Average number of steps") + theme_bw()
```

